---
title: "ProjectMachineLearning"
author: "zoran"
date: "22 March 2015"
output: pdf_document
---

# Overview

This project is focused on using machine learning algorithms to predict the manner in which 6 participants have performed weight lifting exercise, in terms of 'how well' they are doing it - as per criteria described in the paper in http://groupware.les.inf.puc-rio.br/har. This site also containts data that are used for training and testing.

# Detailed description

## Data cleaning

We have first read the .csv data and cleaned them, in particular removing columns that have NA charaters, as shown below

```{r}
library(caret)
library(randomForest)
# read CSV data first - and define a character string c with all "NA" characters 
trainingdata <- read.csv("./pml-training.csv",na.strings=c("NA","#DIV/0!",""))
testingdata <- read.csv("./pml-testing.csv",na.strings=c("NA","#DIV/0!",""))
dim(trainingdata); dim(testingdata)

# Clean raw training and testingdata by removing all columns which contain all NA data
training <- trainingdata[,colSums(is.na(testingdata))==0]
testing <- testingdata[,colSums(is.na(testingdata))==0]

# Delete first six non activity data
training  <-training[,-c(1:7)]
testing <-testing[,-c(1:7)]
```

## Preparing training and testing sets

The trianing and testing data are created from the cleaned data above.

```{r}
# Split the training data set in training and testing again 
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
subTraining <- training[inTrain,]
subTesting <- training[-inTrain,]
```

## Method used

We used random forest method as it was mentiioned in the paper. We used the randomForest package rather than train caret package, as we found it to perform much faster that the train caret package. 

```{r}
# apply random forest algorithm
modFit <- randomForest(classe ~., data=subTraining, na.action=na.omit)
pred <- predict(modFit, subTesting, type ="class")
```

```{r}
# Check confusion matrix 
confMatrix <- confusionMatrix(pred,subTesting$classe)
confMatrix
```

```{r]}
## create anwsers 
answers <- predict(modFit, testing, type="class")
anwsers
```

The following is function recommended in the assignment for submitting results

```{r}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(answers)
```
